---
title: "Final"
author: "Apurva Hari"
date: "August 23, 2016"
output: pdf_document

---

#### Consider Anscombe's Quartet, a dataset included with R, designed by a mathematician to demonstrate the dangers in relying only on statistical measurements of a rstdataset

```{r}
#reading the dataset Anscombe
data_set <- anscombe
head(data_set)

#reshaping the data so that each (x,y) pair are displayed together
mydata=with(anscombe,data.frame(xVal=c(x1,x2,x3,x4), yVal=c(y1,y2,y3,y4), group=gl(4,nrow(anscombe))))

#Finding out the mean and standard deviation of each (x,y) pair
aggregate(.~group,data=mydata,mean)
aggregate(.~group,data=mydata,sd)

#Finding out the correlation of each quartet i.e. (x,y) pair
sapply(1:4, function(x) cor(anscombe[, x], anscombe[, x+4]))

#Applying a linear model on each quartet
lm1 <- lm(y1 ~ x1, data = anscombe)
lm2 <- lm(y2 ~ x2, data = anscombe)
lm3 <- lm(y3 ~ x3, data = anscombe)
lm4 <- lm(y4 ~ x4, data = anscombe)

#Finding the summary of each linear model
summary(lm1)
summary(lm2)
summary(lm3)
summary(lm4)

#Plotting each of the (x,y) pairs separately and fitting a line through each quartet
library(ggplot2)
library(gridExtra)

plt1 <- ggplot(anscombe) + geom_point(aes(x1, y1), color = "red", size = 3) + scale_x_continuous(breaks = seq(0, 20, 2)) + scale_y_continuous(breaks = seq(0, 12, 2)) + geom_abline(intercept = 3, slope = 0.5, color = "green") + expand_limits(x = 0, y = 0) + labs(title = "dataset 1")

plt2 <- ggplot(anscombe) + geom_point(aes(x2, y2), color = "red", size = 3) + scale_x_continuous(breaks = seq(0, 20, 2)) + scale_y_continuous(breaks = seq(0, 12, 2)) + geom_abline(intercept = 3, slope = 0.5, color = "green") + expand_limits(x = 0, y = 0) + labs(title = "dataset 2")

plt3 <- ggplot(anscombe) + geom_point(aes(x3, y3), color = "red", size = 3) + scale_x_continuous(breaks = seq(0, 20, 2)) + scale_y_continuous(breaks = seq(0, 12, 2)) + geom_abline(intercept = 3, slope = 0.5, color = "green") + expand_limits(x = 0, y = 0) + labs(title = "dataset 3")

plt4 <- ggplot(anscombe) + geom_point(aes(x4, y4), color = "red", size = 3) + scale_x_continuous(breaks = seq(0, 20, 2)) + scale_y_continuous(breaks = seq(0, 12, 2)) + geom_abline(intercept = 3, slope = 0.5, color = "green") + expand_limits(x = 0, y = 0) + labs(title = "dataset 4")

grid.arrange(plt1, plt2, plt3, plt4, top = "Anscombe's Quartet")
```

##### data set 1 is clearly linear with some scatter 
##### data set 2 is clearly quadratic  
##### data set 3 has an "outlier" 
##### data set 4 -is not linear at all.



#### Find a dataset pertinent to your work and/or interests, which will be suitable for analysis by ONE of the following:  linear regression

#### Response -I chose to work on the iris dataset. 
#### This famous (Fisher's or Anderson's) iris data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris.The species are Iris setosa, versicolor, and virginica. This is a very well-known dataset and is widely used. I chose this, as it has well established data that will help me understand how to explore data and also fit a linear model for it. 



####Data Exploration of the Iris dataset.Execute the chosen analysis:   
####Linear simple or multiple regression:  fitting model of all or chosen predictors, maybe some with interactions, maybe using quadratic or log transformations, if appropriate


``` {r}

#Checking the dimensionality
dim(iris)
#Structure of iris
str(iris)
#getting the first 5 rows
iris[1:5,]
#Distribution of every variable
summary(iris)
#Frequency
table(iris$Species)
#covariance of 2 variables
cov(iris$Petal.Length,iris$Sepal.Length)
#correlation between 2 variables
cor(iris$Sepal.Length,iris$Petal.Length)
#mean of all the flowers
aggregate(iris[,1:4], by=list("Species" = iris$Species), mean)


#pie chart
pie(table(iris$Species))
#Histogram
hist(iris$Petal.Length)
#density
plot(density(iris$Sepal.Length))
#Scatter plot
plot(iris$Petal.Length,iris$Petal.Width)
#Pair Plot
pairs(iris)
```
#### Execute the chosen analysis:
#### Linear simple or multiple regression-fitting model of all or chosen predictors, maybe some with interactions, maybe using quadratic or log transformations, if appropriate. Evaluate your results, and possibly try some and evaluate minor refinements.  

```{r}

#Trying a linear regression to predict the petal width
fit <- lm(Petal.Width ~ Petal.Length, data=iris)
class(fit)
summary(fit)

coefficients(fit) # model coefficients
predict(fit) # fitted predictions
predict(fit, newdata=data.frame(Petal.Length=seq(1, 2, by=0.1)))
confint(fit, level=0.95) # Confidence Intervals for model parameters 
fitted(fit) # predicted values
residuals(fit) # residuals
influence(fit) # regression diagnostics

par(mfrow=c(2,2))
plot(fit)


#Trying a Multiple Linear Regression
fit2 <- lm(Petal.Width ~ Petal.Length + Sepal.Length + Sepal.Width, data=iris)
summary(fit2) # show results

#Finding the Interaction Terms and do a linear fit

fit2int <- lm(Petal.Width ~ Petal.Length + Sepal.Length + Sepal.Width + Petal.Length:Sepal.Length, data=iris)

#Analyzing covariance using linear regression
fit3 <- lm(Petal.Width ~ Petal.Length + Sepal.Length + Sepal.Width + Species, data=iris)
summary(fit3)
```

#### Evaluate your results, and possibly try some and evaluate minor refinements.  Use graphs of the model (Kabacoff, Sec. 8.3.2),  assess basic assumptions like , normality, independence, linearity, homoscedasticity; gvlma() in package "gvlma";  look for and possibly deal with outliers, high-leverage points, influential observations.  
```{r}
#Evaluating regression using plots
library(car)
#Checking the Homoscedasticity of the fitted model
ncvTest(fit)
spreadLevelPlot(fit) #creating a scatter-plot

#Using crplots to check for linearity
crPlots(fit)
crPlots(fit3)

#Global validation of the linear model assumption
library(gvlma)

gvlmodel <- gvlma(fit)
summary(gvlmodel)


gvlmodel2 <- gvlma(fit3)
summary(gvlmodel2)

#Checking Outliers for a few of the linear models

outlierTest(fit)

outlierTest(fit2)

outlierTest(fit3)

```





